---
title: "ungulates"
author: "Ilya"
date: "12/7/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####install packages
```{r packages, echo=FALSE}
#install.packages("rlang")
#library("rlang")
pkgTest <- function(x)
{
  if (x %in% rownames(installed.packages()) == FALSE) {
    install.packages(x, dependencies= TRUE)    
  }
  library(x, character.only = TRUE)
}
neededPackages <- c("leaflet", "raster", "mapview", "sp", "ggmap",
                    "rasterVis",   "ggplot2", "dplyr", "RgoogleMaps", "rworldmap", "lubridate", "zoo", "broom")

for (package in neededPackages){pkgTest(package)}

```

###summarize EMPRES-i data
```{r empresi}
E = read.csv("Outbreak_list4775267628333941320.csv")
dim(E)
summary(E)
save(E, file = "E.Rdata")
df = E
df$logsumCases = log(df$sumCases+0.01)
#df = subset(df, !is.na(sumCases))
proj  = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# following directions here to make dataframe spatial: https://stackoverflow.com/questions/29736577/how-to-convert-data-frame-to-spatial-coordinate"

xy <- cbind(df$longitude,df$latitude)#package sp

df <-SpatialPointsDataFrame(coords = xy, data = df, 
                            proj4string = CRS(proj))

plot <- ggplot(E, aes(x = longitude, y = latitude))+
  geom_point(alpha = 0.1)

plot
ggsave(plot = plot, file = "empres_disease.jpg")

size = 10
plot <- ggplot(E, aes(x = disease))+
  geom_histogram(stat = "count")+
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size =size))

plot
ggsave(plot = plot, file = "empres_disease_hist.jpg")

# pal <- colorNumeric("viridis", NULL)
#commenting this out because it does not display well in github_document
# M<- leaflet(df) %>%
#   addTiles() %>%
#   addCircleMarkers(color = ~pal(disease),
#     stroke = FALSE, fillOpacity = 0.1) %>%
#   # setView(lat = 39.5, lng=-98.5, zoom =4) %>%
#    #  add(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
#    #  fillColor = ~pal(outC$incidence),
#    #  #label with county name 
#    #  label = ~paste0(county_name, ": ", formatC(outC$incidence, big.mark = ","))) %>%
#    addLegend(pal = pal, values = df$sumCases, opacity = 1.0,
#      labFormat = labelFormat(transform = function(x) round(10^x)))
# 
# mapshot(M, file = "EMPRESi_cases.png")
#M
```

####Parse species field
Make separate row for each animal host species. Separate by commas. Create fields animal.origin (wild vs. domestic) and animal.species. Save out.Rdata
```{r parse_species}
source("parse_species.R")
```

####Graph locations of wild vs. domestic species
```{r graph_wild_domestic_location}
source("graph_wild_domestic_location.R")
```

####Subset to only include mammalian livestock. Save out_mammals.Rdata
```{r  subset_mammals}
source("subset_mammals.R")
```

####Graph locations of each mammal livestock species
```{r graph_mammal_livestock}
source("graph_mammal_livestock.R")
```

####Graph locations of each disease
```{r  graph_disease_map}
source("graph_disease_map.R")
```



####Read in FAO/ILRI livestock production zones map (source:  https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/WPDSZE/GQAXNS&version=1.1). Generate and sample random points. Limit latitude range of random points to y<80 & y >-60 so that they fall within deforestation tiles. Exclude random points that lie within areas deemed unsuitable by FAO. Save r_df_sample.Rdata
```{r livestock_production_systems}
source("livestock_production_systems.R")
```

####Fix time
```{r time}
source("time.R")
```

####Graph outbreaks over time in livestock and wildlife
```{r graph_time}
source("graph_time.R")
# source("graph_time_datetime.R")
source("graph_time_datemonth.R")
```


####Graph outbreaks over time, stacked by disease
```{r}
source("graph_time_diseases.R")
```

####Graph outbreaks over time, one facet per disease
```{r}
source("graph_time_diseases_facet.R")
```



####Combine presences and absences. Use "out_mammals.Rdata" and "r_df_sample.Rdata". Output: df_combined.Rdata
```{r presence_absence_combine}
source("presence_absence_combine.R")

```

###Assign environmental data to presence-absence data

####Deforestation
global forest change
source: https://earthenginepartners.appspot.com/science-2013-global-forest/download_v1.5.html
```{r deforest_sample}
#Year of gross forest cover loss event (lossyear)
#Forest loss during the period 2000–2017, defined as a stand-replacement disturbance, or a change from a forest to non-forest state. Encoded as either 0 (no loss) or else a value in the range 1–17, representing loss detected primarily in the year 2001–2017, respectively.
#Granule with top-left corner at 10N, 20E (central Africa)
r = raster("Hansen_GFC-2017-v1.5_lossyear_10N_020E.tif")
plot(r)
```

Add fields to df_combined latitude_10 (nearest top 10 degree) and longitude_10 (left 10 degree). Resave df_combined.Rdata
```{r deforestation}
source("deforestation.R")
```

For each 10X10 degree quadrant of outbreak presence / absence data, download relevant tile of deforestation data, and extract deforestation data. Use as input df_combined.Rdata. Save out_deforest.  
```{r getdef}
source("getdef.R")
```

Add field indicating whether location has been deforested or not. Input out_deforest. Make graphs of deforest vs. no for outbreak vs. background locations. Input "out_deforest.Rdata"
```{r deforest_binary}
source("deforest_binary.R")
```


####Graph locations of outbreaks involving wild only, domestic only, or both
```{r graph_origin_construct}
source("graph_origin_construct.R")
```



###Gridded Livestock of the World
####Cattle
```{r cattle}
# source("cattle.R")
```

###History Database of the Global Environment
rangeland, 2017
```{r history_plot}
source("history_plot.R")
```

converted rangeland, 2017
```{r history_plot_conv_rangeland_2017}
source("history_plot_conv_rangeland_2017.R")
```


pasture, 2017
```{r history_plot_pasture_2017}
#source("history_plot_pasture_2017.R")
```

####after this proceed to ungulates1.Rmd


