---
title: "ungulates4"
author: "Ilya"
date: "3/4/2019"
output: github_document
---
##to change in version on workstation 3/15:
###plot predictions
###sdf = dfPredictors


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####install packages
```{r packages, echo=FALSE}
#install.packages("rlang")
#library("rlang")
pkgTest <- function(x)
{
  if (x %in% rownames(installed.packages()) == FALSE) {
    install.packages(x, dependencies= TRUE)    
  }
  library(x, character.only = TRUE)
}
neededPackages <- c("leaflet", "raster", "mapview", "sp", "ggmap",
                    "rasterVis",   "ggplot2", "dplyr", "RgoogleMaps", "rworldmap", "lubridate", "zoo", "broom", "reticulate", "gbm", "caret", "caTools", "ROCR", "grDevices")

for (package in neededPackages){pkgTest(package)}
```

###load data on outbreaks and on road density
```{r road_density}
load("dfAnimal.Rdata")
df = dfAnimal
out_spdf = SpatialPointsDataFrame(coords = df[,c("longitude", "latitude")], 
                               data = df)

###road density
R = raster("DATA/GRIP4_density_total/grip4_total_dens_m_km2.asc")

#assign projection
projection(R) <-CRS("+proj=longlat +datum=WGS84")
projection(out_spdf) = projection(R)

locs.vals = raster::extract(R, out_spdf,
                            df = TRUE, method = "simple")#specify 
#make into dataframe
locs.vals.df = data.frame(locs.vals)
#create row indices
locs.vals.df$row = seq(1, dim(locs.vals.df)[1])
# names(locs.vals.df)[2]="globcover"
df$row = seq(1, dim(locs.vals.df)[1])

df= merge(df, locs.vals.df, by = "row")
dfRoad = df
save(dfRoad, file = "dfRoad.Rdata")


```

###add log livestock density 
```{r livestock_density}
load("dfRoad.Rdata")
df = dfRoad
names(df)[names(df) == "DATA/Chickens10km_AD_2010_v2_01"] = "chickens2010"
names(df)[names(df) == "DATA/Sheep10km_AD_2010_v2_1"] = "sheep2010"
names(df)[names(df) == "DATA/Goats10km_AD_2010_v2_1"] = "goats2010"
names(df)[names(df) == "DATA/Pigs10km_AD_2010_GLW2_01"] = "pigs2010"
names(df)[names(df) == "DATA/Cattle10km_AD_2010_v2_1"] = "cattle2010"

addSmall = 0.0001
df$log_chickens = log(df$chickens2010+addSmall)
df$log_sheep = log(df$sheep2010+addSmall)
df$log_goats = log(df$goats2010+addSmall)
df$log_cattle = log(df$cattle2010+addSmall)
df$log_pigs = log(df$pigs2010+addSmall)

#add column indicating number of livestock species with nonzero value
for (a in 1:dim(df)[1]){
  animals_tmp = c(df$chickens2010[a],
                  df$pigs2010[a],
                  df$goats2010[a],
                  df$sheep2010[a],
                  df$cattle2010[a])
  df$livestock_non_zero_fraction[a] = length(which(animals_tmp>0))/
    length(animals_tmp)
}
dfAnimalRoad = df
save(dfAnimalRoad, file = "dfAnimalRoad.Rdata")
```

###add data from one year of nightlights
```{r nightlights}

load("dfAnimalRoad.Rdata")
df = dfAnimalRoad
out_spdf = SpatialPointsDataFrame(coords = df[,c("longitude", "latitude")], 
                               data = df)

###lights
R = raster("DATA/F182013.v4/F182013.v4c_web.stable_lights.avg_vis.tif")

#assign projection
projection(R) <-CRS("+proj=longlat +datum=WGS84")
projection(out_spdf) = projection(R)

locs.vals = raster::extract(R, out_spdf,
                            df = TRUE, method = "simple")#specify 
#make into dataframe
locs.vals.df = data.frame(locs.vals)
#create row indices
locs.vals.df$row = seq(1, dim(locs.vals.df)[1])
# names(locs.vals.df)[2]="globcover"
df$row = seq(1, dim(locs.vals.df)[1])

df= merge(df, locs.vals.df, by = "row")
inds255 = which(df$F182013.v4c_web.stable_lights.avg_vis==255)
df$F182013.v4c_web.stable_lights.avg_vis[inds255]=NA#255 is the no-cloudless-sky data value
dfAnimalRoadLight = df
save(dfAnimalRoadLight, file = "dfAnimalRoadLight.Rdata")

```

###get data on outbreaks, background and specific types of road density
```{r roads}
load("dfAnimalRoadLight.Rdata")

df = dfAnimalRoadLight
out_spdf = SpatialPointsDataFrame(coords = df[,c("longitude", "latitude")], 
                               data = df)

#download road density datasets:
#https://www.globio.info/download-grip-dataset
###road density -- type 1, highway
R1 = raster("DATA/GRIP4_density_tp1/grip4_tp1_dens_m_km2.asc")
###road density -- type 2, primary
R2 = raster("DATA/GRIP4_density_tp2/grip4_tp2_dens_m_km2.asc")
roads = addLayer(R1, R2)
###road density -- type 3, secondary
R3 = raster("DATA/GRIP4_density_tp3/grip4_tp3_dens_m_km2.asc")
roads = addLayer(roads, R3)
###road density -- type 4, tertiary
R4 = raster("DATA/GRIP4_density_tp4/grip4_tp4_dens_m_km2.asc")
roads = addLayer(roads, R4)
###road density -- type 5, local
R5 = raster("DATA/GRIP4_density_tp5/grip4_tp5_dens_m_km2.asc")
roads = addLayer(roads, R5)
names(roads)=c("highway", "primary_roads", "secondary_roads",
               "tertiary_roads", "local_roads")
#assign projection
projection(roads) <-CRS("+proj=longlat +datum=WGS84")
projection(out_spdf) = projection(roads)

locs.vals = raster::extract(roads, out_spdf,
                            df = TRUE, method = "simple")#specify 
#make into dataframe
locs.vals.df = data.frame(locs.vals)
#create row indices
locs.vals.df$row = seq(1, dim(locs.vals.df)[1])
# names(locs.vals.df)[2]="globcover"
df$row = seq(1, dim(locs.vals.df)[1])

df= merge(df, locs.vals.df, by = "row")

addSmall = 0.0001
df$log_highway_density = log(df$highway+addSmall)
df$log_primary_road_density = log(df$primary_roads+addSmall)
df$log_secondary_road_density = log(df$secondary_roads+addSmall)
df$log_tertiary_road_density = log(df$tertiary_roads+addSmall)
df$log_local_road_density = log(df$local_roads+addSmall)
dfRoads = df

save(dfRoads, file = "dfRoads.Rdata")


```


###assign land cover to outbreak and background points. output: 
#dfRoads_LC.Rdata
```{r lc}
source("R_landcover_assign.R")
```

###hot one encode land cover
```{r lc_one_hot}
load("dfRoads_LC.Rdata")
df = dfRoads_LC
inds = c(5:6,31,50:57,64:68,70)
df = df[,inds]
#Converting every categorical variable to numerical using dummy variables
dmy <- dummyVars(" ~ .", data = df,fullRank = T, sep=" ")
df_transformed <- data.frame(predict(dmy, newdata = df))
dfLC_onehot = df_transformed
save(dfLC_onehot, file = "dfLC_onehot.Rdata")
```

##assign worldclim climate variables to outbreak and background points
```{r worldclim}
load("dfLC_onehot.Rdata")
df = dfLC_onehot

#http://worldclim.org/version2
path <- "DATA/wc2.0_30s_bio/"

grids <- list.files(path , pattern = "*.tif$")

#create a raster stack from the input raster files 
w <- stack(paste0(path, grids))

w <- subset(w, c("wc2.0_bio_30s_01", "wc2.0_bio_30s_12"))
names(w) = c("annualMeanTemp", "annualPrecip")

out_spdf = SpatialPointsDataFrame(coords = df[,c("longitude", "latitude")], 
#assign projection
                               data = df)
projection(w) <-CRS("+proj=longlat +datum=WGS84")
projection(out_spdf) = projection(w)

locs.vals = raster::extract(w, out_spdf,
                            df = TRUE, method = "simple")#specify 
#make into dataframe
locs.vals.df = data.frame(locs.vals)
#create row indices
locs.vals.df$row = seq(1, dim(locs.vals.df)[1])
# names(locs.vals.df)[2]="globcover"
df$row = seq(1, dim(locs.vals.df)[1])

df= merge(df, locs.vals.df, by = "row")

save(df, file = "df.Rdata")

```

###assign country from shapefile to outbreak and background
```{r country}
load("df.Rdata")
r = shapefile("DATA/TM_WORLD_BORDERS-0.3/TM_WORLD_BORDERS-0.3.shp")

out_spdf = SpatialPointsDataFrame(coords = df[,c("longitude", "latitude")], 
#assign projection
                               data = df)
projection(r) <-CRS("+proj=longlat +datum=WGS84")
projection(out_spdf) = projection(r)
out_spdf$country <- over(out_spdf, r)$NAME#name of park
#make into dataframe
df= as.data.frame(out_spdf)
keep = names(df)
rm_fields = c("latitude.1", "longitude.1", "ID")
keep = setdiff(keep, rm_fields)
df = df[,keep]

#change to one-hot encoding of country
#Converting every categorical variable to numerical using dummy variables
dmy <- dummyVars(" ~ .", data = df,fullRank = T, sep=" ")
df_transformed <- data.frame(predict(dmy, newdata = df))
df = df_transformed

save(df, file = "df.Rdata")
```

###construct interaction features in df: log of the product of roadDensity and livestock density 
```{r interaction}
load("df.Rdata")
df$log_road_chickens = log(df$grip4_total_dens_m_km2*exp(df$log_chickens))
df$log_road_pigs = log(df$grip4_total_dens_m_km2*exp(df$log_pigs))
df$log_road_cattle = log(df$grip4_total_dens_m_km2*exp(df$log_cattle))
df$log_road_sheep = log(df$grip4_total_dens_m_km2*exp(df$log_sheep))
df$log_road_goats = log(df$grip4_total_dens_m_km2*exp(df$log_goats))
save(df, file = "df.Rdata")
```

###get predictors for outbreak and background for gbm-- log road density w/ one hot land cover, climate, and country
```{r gbm_predictors}
load("df.Rdata")
names(df)[names(df)== "F182013.v4c_web.stable_lights.avg_vis"]="nightlights2013"
names(df)[names(df)== "grip4_total_dens_m_km2"]="roadDensity"
df$logRoadDensity = log(df$roadDensity+0.0001)

rm_fields = c("roadDensity", "latitude", "longitude", "row")
okay_fields = setdiff(names(df), rm_fields)
df = df[,okay_fields]
dfPredictors = df
save(dfPredictors, file = "dfPredictors.Rdata")

model<-as.formula(paste(colnames(df)[1], "~",
                        paste(colnames(df)[c(2:230)],collapse = "+"),
                        sep = ""))
print(model)
save(model, file = "model.Rdata")

```

###partition train and test -- log road density -- multiple road types -- land cover -- climate -- country
```{r partition}
library("caret")
load("dfPredictors.Rdata")
df = dfPredictors
DP =createDataPartition(y = df$case, 
                        p = 0.8,
                        list = FALSE)
Train = df[DP,]
summary(Train$case)
Test = df[-DP,]

save(Train, file = "Train.Rdata")
#save(Test, file = "Test.Rdata")

```


###run GBM -- predictors include roads, livestock, nightlights (2013), land cover, climate, country
```{r gbm_run}

load("Train.Rdata")
# #https://stackoverflow.com/questions/19463137/error-in-evalexpr-envir-enclos-object-not-found
attach(Train)
load("model.Rdata")
#Start the clock
ptm<-proc.time()

n.trees = 10000
shrinkage = 0.001#final version should be 0.001
cv.folds = 10#final version should be 10
gbmtest<- gbm(model,
              data=Train,
              distribution="bernoulli",
              n.trees=n.trees,
              shrinkage=shrinkage,
              interaction.depth=4,
              bag.fraction=0.50,
              train.fraction=1,
              n.minobsinnode=5,
              cv.folds=cv.folds,
              keep.data=TRUE,
              verbose=TRUE,
              n.cores=NULL)

save(gbmtest, file = "gbmtest.Rdata")
#check performance using 5-fold cross-validation
best.iter <- gbm.perf(gbmtest,method="cv",plot.it=FALSE) #this gives you the optimal number of trees based on cv performance, other methods will over or under predict
print(best.iter)

gbm_error = data.frame(train.error = gbmtest$train.error,
                       trees = seq(1,n.trees))
plot <- ggplot(gbm_error, aes(x = trees, y = train.error))+
  geom_line()
plot
ggsave(filename = "deviance_RoadAnimalNightLcClimateCountry.jpg",
       plot = plot)
#Stop the clock
(proc.time()-ptm)/60

# output predictions on the TRAINING SET
output<-predict(gbmtest,
                newdata=Train,
                n.trees=best.iter,
                type="response")

#need to figure out how to interpret this, or change it to be interpretable
output<-cbind(output,Train$case)
colnames(output)<-c("output","data")
rownames(output)<-rownames(Train)
output<-output[order(-output[,1]),]

# # AUC for Bernoulli distributed responses
par(mar = c(1,1,1,1))
auc=colAUC(output[,1],output[,2],
           plotROC = TRUE)

print(auc)
pred<-prediction(output[,1],output[,2])
perf<-performance(pred,"tpr","fpr")

par(mar = c(1,1,1,1))
plot(perf,colorize=TRUE,main="ROC full model")
abline(a=0, b= 1)

```

###plot relative influence -- multiple road types
###version with cutoff threshold of minimum interesting influence of 1%
```{r rel_influence_1}
#format relative influence for figure
load("gbmtest.Rdata")
# gbmtest = gbmtestRoadsAnimalNight
x = summary(gbmtest)
# 
x.df= data.frame(variable = x$var,
                 relative.influence = x$rel.inf)

x.df = subset(x.df, relative.influence >=1)

x.df$variable = factor(x.df$variable, levels = x.df$variable[order(x.df$relative.influence)])
save(x.df, file = "x.df.Rdata")
ggplot(data = x.df, aes(x = variable, y =relative.influence))+
  ylab("relative influence (%)")+
  xlab("variable")+
  geom_bar(stat="identity")+
  coord_flip()
# 
ggsave("Figure.relative.influence.Animals.Roads.Climate.Nightlights.Country.jpg")

```

###partial dependency plot
Get data needed for partial dependency plot -- multiple road types. Requires gbmtest and Train. Makes "out_partial.Rdata"
```{r partial_dependence1}
load("gbmtest.Rdata")
load("x.df.Rdata")
# gbmtest = gbmtest
#https://stackoverflow.com/questions/26917689/how-to-use-facets-with-a-dual-y-axis-ggplot

# Data
out = NULL
#x = summary(gbmtest)
#varlist = x$var
#get variables from gbmtest
#varlist = gbmtest$var.names
#get only those variables that have relative influence above certain threshold
varlist = x.df$variable
for (i in 1:length(varlist)){#begin for loop
  #print(i)
  i.var = which(gbmtest$var.names==varlist[i])
  plot.out = plot.gbm(gbmtest, i.var = as.character(varlist[i]), return.grid = TRUE)
  names(plot.out)[1]="variable.value"
  names(plot.out)[2]="value"
  plot.out$variable.name=varlist[i]
  plot.out$var = "marginal.effect"#for plotting
  out = rbind(out, plot.out)
}#end for loop

Train1=subset(Train, case == 1)

i =1
i = 3
for (i in 1:length(varlist)){#begin for loop through variables
  print(i)
  i.var = which(names(Train1)==varlist[i])
  h = hist(Train1[,i.var], plot = TRUE)
  tmp = data.frame(variable.value = h$mids,
                   value=h$counts/sum(h$counts),#normalize
                   variable.name=varlist[i],
                   var = "frequency")
  out = rbind(out, tmp)
}#end for loop
out_partial = out
save(out_partial, file = "out_partial.Rdata")

```


###Make partial dependency plots
```{r partial_dependence_plot}
library(latticeExtra)
#use latticeExtra to make two plots
load("x.df.Rdata")
load("out_partial.Rdata")
#https://stat.ethz.ch/R-manual/R-devel/library/grDevices/html/png.html
jpeg(filename = "Figure.partial.dependency.Night.Animal.Log.Roads.jpeg", width = 960, height = 960, quality = 95)
out = out_partial
x.df.sorted = sort(x.df$relative.influence, decreasing = FALSE, index.return= TRUE)
x.df$sort_index_decreasing = x.df.sorted$ix
x.df = subset(x.df, relative.influence>0)

var.plot = x.df$variable

out$variable.name=as.character(out$variable.name)

neworder <- x.df$variable#var.plot
library(plyr)  ## or dplyr (transform -> mutate)
dat <- arrange(transform(out,
                         variable.name=factor(variable.name,levels=neworder)),variable.name)
out = dat

## a variant of Figure 5.13 from Sarkar (2008)
## http://lmdvr.r-forge.r-project.org/figures/figures.html?chapter=05;figure=05_1

x_between = 5#not sure what this does, seems to affect size of subplot
x_axis_cex = 1
names(out)
head(out)
out$value = round(out$value, digits = 1)
out_marg_eff = subset(out, var == "marginal.effect")
out_marg_eff$value=round(out_marg_eff$value, digits = 1)
marg_eff <- xyplot(value ~ variable.value | variable.name,
                   data = out_marg_eff, type = "l", 
                   #layout = c(4, 3),
                   scales = list(relation = "free", x=list(cex=x_axis_cex)),
                   between = list(x = x_between),
                   ylab = "Marginal effect",
                   xlab = "Predictor value",
                   auto.key = FALSE,#legend,
                   par.settings = list(strip.background=list(col="lightgrey")),
                   par.strip.text=list(cex=1),
                   as.table = TRUE
)

out_count = subset(out, var == "frequency")
count_plot <- xyplot(value ~ variable.value | variable.name, data = out_count, type = "h",
                     between = list(x = x_between),
                     scales = list(relation = "free", x=list(cex=x_axis_cex)),
                     ylab = "frequency",
                     xlab = "predictor value",
                     #lattice.options = ggplot2like.opts(),
                     #list(superpose.symbol = list(col = c("blue")))
                     # par.settings = ggplot2like(),
                     auto.key=FALSE,
                     as.table = TRUE
                     
)
#count_plot

# doubleYScale(marg_eff, count_plot, style1 = 0, style2 = 3, add.ylab2 = TRUE,
#    text = c("marginal effect", "frequency"), columns = 2)

plot <- doubleYScale(marg_eff, count_plot, style1 = 0, style2 = 3, add.ylab2 = TRUE, columns = 2)
plot
dev.off()

```




###make correlogram
```{r correlogram}
library(corrplot)
#load("dfPredictorsRoadsAnimalNight.Rdata")
load("dfPredictors.Rdata")
M = cor(df, use = "complete.obs")
corrplot(M, method = "square",
         tl.cex=0.5,
         type = "lower")

```




#make predictions
###get land cover and add to other predictor fields
###test out  assembling predictors
```{r predictors_test}
#livestock density from 2010

# pig2010 = raster("Pigs10km_AD_2010_GLW2_01_TIF/Pigs10km_AD_2010_GLW2_01.tif")
# 
# cattle2010 = raster("Cattle10km_AD_2010_GLW2_01/Cattle10km_AD_2010_v2_1.tif")
# 
# animal2010 = addLayer(cattle2010, pig2010)
# 
# #assign a projection
# projection(animal2010) <-CRS("+proj=longlat +datum=WGS84")
# 
# #convert raster to points (package raster)
# predictorPointsCattlePig = rasterToPoints(animal2010, spatial=TRUE)
# 
# summary(predictorPointsCattlePig)
# 
# #now try converting cattle without adding layer
# tmp = rasterToPoints(cattle2010, spatial = TRUE)
# dim(tmp)
# #summary(Cattle)
# 
# out = NULL
# out[[1]]=data.frame(tmp)
# 
# tmp = rasterToPoints(pig2010, spatial = TRUE)
# dim(tmp)
# out[[2]]=data.frame(tmp)
# 
# out2= NULL
# out2 = out[[1]]
# for (a in 2){
#   out2 = cbind(out2, out[[a]])
# }

```


###assemble predictors into one raster brick
```{r predictors_brick}

path = "DATA/2_GlobalRuminantLPS_GIS/"
r  = raster(paste0(path,"glps_gleam_61113_10km.tif"))

#livestock density from 2010
pig2010 = raster("DATA/Pigs10km_AD_2010_GLW2_01_TIF/Pigs10km_AD_2010_GLW2_01.tif")

cattle2010 = raster("DATA/Cattle10km_AD_2010_GLW2_01/Cattle10km_AD_2010_v2_1.tif")

animal2010 = addLayer(pig2010, cattle2010)

#add layer of land use
animal2010 = addLayer(animal2010, r)

goats2010 = raster("DATA/Goats10km_AD_2010_v2_1_TIF/Goats10km_AD_2010_v2_1.tif")

animal2010 = addLayer(animal2010, goats2010)

sheep2010 = raster("DATA/Sheep10km_AD_2010_GLW2_1_TIF/Sheep10km_AD_2010_v2_1.tif")

animal2010 = addLayer(animal2010, sheep2010)

chickens2010 = raster("DATA/Chickens10km_AD_2010_v2_01_TIF/Chickens10km_AD_2010_v2_01.tif")

animal2010 = addLayer(animal2010, chickens2010)
names(animal2010)= c("pigs2010", "cattle2010", "livestock_system", "goats2010", "sheep2010", "chickens2010")
R = animal2010
dim(R)

#https://www.globio.info/download-grip-dataset
###road density -- type 1, highway
R1 = raster("DATA/GRIP4_density_tp1/grip4_tp1_dens_m_km2.asc")
###road density -- type 2, primary
R2 = raster("DATA/GRIP4_density_tp2/grip4_tp2_dens_m_km2.asc")
roads = addLayer(R1, R2)
###road density -- type 3, secondary
R3 = raster("DATA/GRIP4_density_tp3/grip4_tp3_dens_m_km2.asc")
roads = addLayer(roads, R3)
###road density -- type 4, tertiary
R4 = raster("DATA/GRIP4_density_tp4/grip4_tp4_dens_m_km2.asc")
roads = addLayer(roads, R4)
###road density -- type 5, local
R5 = raster("DATA/GRIP4_density_tp5/grip4_tp5_dens_m_km2.asc")
roads = addLayer(roads, R5)

###road density -- total
R_road = raster("DATA/GRIP4_density_total/grip4_total_dens_m_km2.asc")
roads = addLayer(roads, R_road)
names(roads)=c("highway", "primary_roads", "secondary_roads",
               "tertiary_roads", "local_roads", "RoadDensity")
dim(roads)

R = addLayer(R, roads)

#assign a projection
projection(R) <-CRS("+proj=longlat +datum=WGS84")

Gl = raster("DATA/Globcover2009_V2.3_Global_/GLOBCOVER_L4_200901_200912_V2.3.tif")
projection(Gl) <- CRS("+proj=longlat +datum=WGS84")

#x = object to be resampled
#y = object to be resampled to
#resample with nearest neighbor for categorical data 
#https://gis.stackexchange.com/questions/314960/resampling-raster-with-categorical-data-in-r
Gl_resampled = resample(x = Gl, y = R, method = "ngb")
save(Gl_resampled, file = "Gl_resampled.Rdata")
R = addLayer(R, Gl_resampled)
R_nightlights = raster("DATA/F182013.v4/F182013.v4c_web.stable_lights.avg_vis.tif")
dim(R_nightlights)
names(R_nightlights)= "nightlights2013"#give it name that matches analysis name
#x = object to be resampled
#y = object to be resampled to
R_nightlights_resampled = resample(x = R_nightlights, y = R)

R = addLayer(R, R_nightlights_resampled)

#################WorldClim
path <- "DATA/wc2.0_30s_bio/"

grids <- list.files(path , pattern = "*.tif$")

#create a raster stack from the input raster files 
w <- stack(paste0(path, grids))

w <- subset(w, c("wc2.0_bio_30s_01", "wc2.0_bio_30s_12"))
names(w) = c("annualMeanTemp", "annualPrecip")
worldclim = w
# save(worldclim, file = "worldclim.Rdata")

w = resample(x = w, y = R)
R = addLayer(R, w)
#################Worldclim

#convert raster to points (package raster)
predictorPoints = rasterToPoints(R, spatial=TRUE)
save(predictorPoints, file = "predictorPoints.Rdata")

```

#assign label_short describing land cover to predictorPoints
```{r label_short_predictors}
load("predictorPoints.Rdata")
G_legend = read.csv("DATA/Globcover2009_V2.3_Global_/Globcover2009_Legend.csv")
names(G_legend)[names(G_legend)=="Value"]="globcover"
names(predictorPoints)[names(predictorPoints)=="GLOBCOVER_L4_200901_200912_V2.3"]="globcover"
G_legend = G_legend[,c("globcover", "label_short")]
dim(predictorPoints)
predictorPoints1 =  merge(predictorPoints, G_legend)
dim(predictorPoints1)
predictorPoints = predictorPoints1
save(predictorPoints, file = "predictorPoints.Rdata")
```

#one-hot encode label_short describing land cover for predictorPoints
```{r lc_one_hot_predictors}
load("predictorPoints.Rdata")
df = predictorPoints
keep = names(df)
keep = setdiff(keep, "globcover")
df = df[,keep]
#Converting every categorical variable to numerical using dummy variables
dmy <- dummyVars(" ~ .", data = df,fullRank = T, sep=" ")
df_transformed <- data.frame(predict(dmy, newdata = df))
df = df_transformed
predictorPoints=df

save(predictorPoints, file = "predictorPoints.Rdata")

```

###create log and fractional fields in global predictorPoints
```{r log_fractional}
load("predictorPoints.Rdata")

df = predictorPoints
addSmall = 0.0001
df$log_chickens = log(df$chickens2010+addSmall)
df$log_sheep = log(df$sheep2010+addSmall)
df$log_goats = log(df$goats2010+addSmall)
df$log_cattle = log(df$cattle2010+addSmall)
df$log_pigs = log(df$pigs2010+addSmall)

#assign nonzero fields
df$pigs_nonzero = 1
df$pigs_nonzero[df$pigs2010==0]=0

df$goats_nonzero = 1
df$goats_nonzero[df$goats2010==0]=0

df$sheep_nonzero = 1
df$sheep_nonzero[df$sheep2010==0]=0

df$cattle_nonzero = 1
df$cattle_nonzero[df$cattle2010==0]=0

df$chickens_nonzero = 1
df$chickens_nonzero[df$chickens2010==0]=0

df$livestock_non_zero_fraction = (df$pigs_nonzero+
                                    df$cattle_nonzero+
                                    df$sheep_nonzero+
                                    df$goats_nonzero+
                                    df$chickens_nonzero)/5

remove_fields = c("chickens2010",
                  "sheep2010",
                  "goats2010",
                  "cattle2010",
                  "pigs2010",
                    "pigs_nonzero",
                    "cattle_nonzero",
                  "sheep_nonzero",
                  "goats_nonzero",
                  "chickens_nonzero")

df$log_highway_density = log(df$highway+addSmall)
df$log_primary_road_density = log(df$primary_roads+addSmall)
df$log_secondary_road_density = log(df$secondary_roads+addSmall)
df$log_tertiary_road_density = log(df$tertiary_roads+addSmall)
df$log_local_road_density = log(df$local_roads+addSmall)
df$logRoadDensity = log(df$RoadDensity+addSmall)

remove_fields = c(remove_fields, "highway",
                  "primary_roads",
                  "secondary_roads",
                  "tertiary_roads",
                  "local_roads",
                  "RoadDensity")

names_df = names(df)
keep_fields = setdiff(names_df, remove_fields)

df = df[,keep_fields]

names(df)
predictorPoints = df
save(predictorPoints, file = "predictorPoints.Rdata")
```

###assign country to predictorPoints
```{r country_predictors}
load("predictorPoints.Rdata")

r = shapefile("DATA/TM_WORLD_BORDERS-0.3/TM_WORLD_BORDERS-0.3.shp")

out_spdf = predictorPoints
coordinates(out_spdf) <- ~ x + y

#predictorPoints is already a SpatialPointsDataFrame 
projection(r) <-CRS("+proj=longlat +datum=WGS84")
projection(out_spdf) = projection(r)
out_spdf$country <- over(out_spdf, r)$NAME#name of park
#make into dataframe
df= as.data.frame(out_spdf)
keep = names(df)
rm_fields = c("latitude.1", "longitude.1", "ID")
keep = setdiff(keep, rm_fields)
df = df[,keep]

#change to one-hot encoding of country
#Converting every categorical variable to numerical using dummy variables
dmy <- dummyVars(" ~ .", data = df,fullRank = T, sep=" ")
df_transformed <- data.frame(predict(dmy, newdata = df))
df = df_transformed

predictorPoints = df
save(predictorPoints, file = "predictorPoints.Rdata")
```

###construct interaction features: log of the product of roadDensity and livestock density 
```{r interaction_predictors}
load("predictorPoints.Rdata")
df = predictorPoints
df$log_road_chickens = log(exp(df$logRoadDensity)*exp(df$log_chickens))
df$log_road_pigs = log(exp(df$logRoadDensity)*exp(df$log_pigs))
df$log_road_cattle = log(exp(df$logRoadDensity)*exp(df$log_cattle))
df$log_road_sheep = log(exp(df$logRoadDensity)*exp(df$log_sheep))
df$log_road_goats = log(exp(df$logRoadDensity)*exp(df$log_goats))
dfPredictors = df
save(dfPredictors, file = "dfPredictors.Rdata")
```

###plot predictions
##need to redo this including land cover
```{r predictions_plot}
load("dfPredictors.Rdata")
load("gbmtest.Rdata")
n.trees = 10000

sdf = dfPredictors
sdf = subset(sdf, !is.na(log_cattle))
dim(sdf)[1]
sdf = subset(sdf, !is.na(logRoadDensity))
sdf = subset(sdf, livestock_system != 15)
dim(sdf)[1]

sdf_pred = predict.gbm(gbmtest, newdata = sdf, n.trees=n.trees, type = "response")
  sdf$predicted= sdf_pred
  pred_gbm_season = sdf
  pred_gbm_season=data.frame(pred_gbm_season)
  plot <- ggplot()+
   geom_tile(data = pred_gbm_season, aes(x = x, y = y, alpha = predicted))
  ggsave("predicted.jpg", plot = plot)
  
pred.xy = pred_gbm_season[,c("x", "y", "predicted")]

out_spdf = SpatialPointsDataFrame(coords = pred.xy[,c("x", "y")], 
                               data = pred.xy)

gridded(out_spdf) <- TRUE
out_spdf$z = out_spdf$predicted
dfr = raster(out_spdf)

writeRaster(dfr, "prediction_no_na.tif", format="GTiff", overwrite=TRUE)


```

###SCRATCH below here.

###format date for python
```{r}

# load("dfRoads.Rdata")
# df = dfRoads
# is.na.date = which(is.na(df$date))
# #fix date that has weird format
# df$date[is.na.date]= as.Date(df$reportingDate[is.na.date], format = "%d/%m/%Y")
# df$month = month(df$date)
# is.na.month = which(is.na(df$month))#should be empty
# 
# #beginning of the year
# df$year_begin = paste0(df$year,"0101")
# #beginning of the previous year
# df$previous_year_begin = paste0(df$year-1,"0101")
# #next year beginning
# df$next_year_begin = paste0(df$year+1,"0101")
# dfDate = df
# save(dfDate, file = "dfDate.Rdata")

```

```{r}
#https://trucvietle.me/r/tutorial/2017/01/18/spatial-heat-map-plotting-using-r.html
# m <- ggplot(faithful, aes(x = eruptions, y = waiting)) +
#  geom_point() +
#  xlim(0.5, 6) +
#  ylim(40, 110)
# m + geom_density_2d()
# m+ stat_density_2d(aes(fill = stat(level)), geom = "polygon")

```

###plot density 
```{r}

# load("dfRoads.Rdata")
# df = dfRoads
# df = df[,c("latitude", "longitude")]
# m <- ggplot(df, aes(x = longitude, y = latitude)) +
#  geom_point() 
#  # xlim(0.5, 6) +
#  # ylim(40, 110)
# m + geom_density_2d()
# m+ stat_density_2d(aes(fill = stat(level)), geom = "polygon")

```

####interpolation with kriging
###https://rpubs.com/nabilabd/118172
```{r}

# load("dfRoads.Rdata")
# df = dfRoads
# library(sp)
# library(gstat)
# suppressPackageStartupMessages({
#   library(dplyr) # for "glimpse"
#   library(ggplot2)
#   library(scales) # for "comma"
#   library(magrittr)
# })
# 
# #1) Converting to an SPDF
# coordinates(df) <- ~ longitude + latitude
# #2) Fitting a variogram
# lzn.vgm <- variogram(case~1, df) # calculates sample variogram values 
# lzn.fit <- fit.variogram(lzn.vgm, model=vgm(1, "Sph", 900, 1)) # fit model

```

