---
title: "ungulates4a livestock predictions"
output: html_document
---


#make predictions
###get land cover and add to other predictor fields
###test out  assembling predictors
```{r predictors_test}
#livestock density from 2010

# pig2010 = raster("Pigs10km_AD_2010_GLW2_01_TIF/Pigs10km_AD_2010_GLW2_01.tif")
# 
# cattle2010 = raster("Cattle10km_AD_2010_GLW2_01/Cattle10km_AD_2010_v2_1.tif")
# 
# animal2010 = addLayer(cattle2010, pig2010)
# 
# #assign a projection
# projection(animal2010) <-CRS("+proj=longlat +datum=WGS84")
# 
# #convert raster to points (package raster)
# predictorPointsCattlePig = rasterToPoints(animal2010, spatial=TRUE)
# 
# summary(predictorPointsCattlePig)
# 
# #now try converting cattle without adding layer
# tmp = rasterToPoints(cattle2010, spatial = TRUE)
# dim(tmp)
# #summary(Cattle)
# 
# out = NULL
# out[[1]]=data.frame(tmp)
# 
# tmp = rasterToPoints(pig2010, spatial = TRUE)
# dim(tmp)
# out[[2]]=data.frame(tmp)
# 
# out2= NULL
# out2 = out[[1]]
# for (a in 2){
#   out2 = cbind(out2, out[[a]])
# }

```


###assemble predictors into one raster brick
```{r predictors_brick}

path = "DATA/2_GlobalRuminantLPS_GIS/"
r  = raster(paste0(path,"glps_gleam_61113_10km.tif"))

#livestock density from 2010
pig2010 = raster("DATA/Pigs10km_AD_2010_GLW2_01_TIF/Pigs10km_AD_2010_GLW2_01.tif")

cattle2010 = raster("DATA/Cattle10km_AD_2010_GLW2_01/Cattle10km_AD_2010_v2_1.tif")

animal2010 = addLayer(pig2010, cattle2010)

#add layer of land use
animal2010 = addLayer(animal2010, r)

goats2010 = raster("DATA/Goats10km_AD_2010_v2_1_TIF/Goats10km_AD_2010_v2_1.tif")

animal2010 = addLayer(animal2010, goats2010)

sheep2010 = raster("DATA/Sheep10km_AD_2010_GLW2_1_TIF/Sheep10km_AD_2010_v2_1.tif")

animal2010 = addLayer(animal2010, sheep2010)

chickens2010 = raster("DATA/Chickens10km_AD_2010_v2_01_TIF/Chickens10km_AD_2010_v2_01.tif")

animal2010 = addLayer(animal2010, chickens2010)
names(animal2010)= c("pigs2010", "cattle2010", "livestock_system", "goats2010", "sheep2010", "chickens2010")
R = animal2010
dim(R)

#https://www.globio.info/download-grip-dataset
###road density -- type 1, highway
R1 = raster("DATA/GRIP4_density_tp1/grip4_tp1_dens_m_km2.asc")
###road density -- type 2, primary
R2 = raster("DATA/GRIP4_density_tp2/grip4_tp2_dens_m_km2.asc")
roads = addLayer(R1, R2)
###road density -- type 3, secondary
R3 = raster("DATA/GRIP4_density_tp3/grip4_tp3_dens_m_km2.asc")
roads = addLayer(roads, R3)
###road density -- type 4, tertiary
R4 = raster("DATA/GRIP4_density_tp4/grip4_tp4_dens_m_km2.asc")
roads = addLayer(roads, R4)
###road density -- type 5, local
R5 = raster("DATA/GRIP4_density_tp5/grip4_tp5_dens_m_km2.asc")
roads = addLayer(roads, R5)

###road density -- total
R_road = raster("DATA/GRIP4_density_total/grip4_total_dens_m_km2.asc")
roads = addLayer(roads, R_road)
names(roads)=c("highway", "primary_roads", "secondary_roads",
               "tertiary_roads", "local_roads", "RoadDensity")
dim(roads)

R = addLayer(R, roads)

#assign a projection
projection(R) <-CRS("+proj=longlat +datum=WGS84")

Gl = raster("DATA/Globcover2009_V2.3_Global_/GLOBCOVER_L4_200901_200912_V2.3.tif")
projection(Gl) <- CRS("+proj=longlat +datum=WGS84")

#x = object to be resampled
#y = object to be resampled to
#resample with nearest neighbor for categorical data 
#https://gis.stackexchange.com/questions/314960/resampling-raster-with-categorical-data-in-r
Gl_resampled = resample(x = Gl, y = R, method = "ngb")
save(Gl_resampled, file = "Gl_resampled.Rdata")
R = addLayer(R, Gl_resampled)
R_nightlights = raster("DATA/F182013.v4/F182013.v4c_web.stable_lights.avg_vis.tif")
dim(R_nightlights)
names(R_nightlights)= "nightlights2013"#give it name that matches analysis name
#x = object to be resampled
#y = object to be resampled to
R_nightlights_resampled = resample(x = R_nightlights, y = R)

R = addLayer(R, R_nightlights_resampled)

#################WorldClim
path <- "DATA/wc2.0_5m_bio/"

grids <- list.files(path , pattern = "*.tif$")

#create a raster stack from the input raster files 
w <- stack(paste0(path, grids))

w <- subset(w, c("wc2.0_bio_5m_01", "wc2.0_bio_5m_12"))
names(w) = c("annualMeanTemp", "annualPrecip")
worldclim = w
# save(worldclim, file = "worldclim.Rdata")

w = resample(x = w, y = R)
R = addLayer(R, w)
#################Worldclim

##########mammal diversity
rMammal = raster("Total mammal richness.gri")
names(rMammal)= "mammalDiversity"

#x = object to be resampled
#y = object to be resampled to
rMammalResampled = resample(x = rMammal, y = R)

R = addLayer(R, rMammalResampled)

##########mammal diversity

#aggregate by factor of five
R <- aggregate(R, fact=5)

#convert raster to points (package raster)
predictorPoints = rasterToPoints(R, spatial=TRUE)
save(predictorPoints, file = "predictorPoints.Rdata")

```

#assign label_short describing land cover to predictorPoints
```{r label_short_predictors}
load("predictorPoints.Rdata")
G_legend = read.csv("DATA/Globcover2009_V2.3_Global_/Globcover2009_Legend.csv")
names(G_legend)[1]="globcover"
names(predictorPoints)[names(predictorPoints)=="GLOBCOVER_L4_200901_200912_V2.3"]="globcover"
G_legend = G_legend[,c("globcover", "label_short")]
dim(predictorPoints)
predictorPoints1 =  merge(predictorPoints, G_legend)
dim(predictorPoints1)
predictorPoints = predictorPoints1
save(predictorPoints, file = "predictorPoints.Rdata")
```

#one-hot encode label_short describing land cover for predictorPoints
```{r lc_one_hot_predictors}
load("predictorPoints.Rdata")
df = predictorPoints
keep = names(df)
keep = setdiff(keep, "globcover")
df = df[,keep]
#Converting every categorical variable to numerical using dummy variables
dmy <- dummyVars(" ~ .", data = df,fullRank = T, sep=".")
df_transformed <- data.frame(predict(dmy, newdata = df))
df = df_transformed
predictorPoints=df

save(predictorPoints, file = "predictorPoints.Rdata")

```

###create log and fractional fields in global predictorPoints
```{r log_fractional}
load("predictorPoints.Rdata")

df = predictorPoints
addSmall = 0.0001
df$log_chickens = log(df$chickens2010+addSmall)
df$log_sheep = log(df$sheep2010+addSmall)
df$log_goats = log(df$goats2010+addSmall)
df$log_cattle = log(df$cattle2010+addSmall)
df$log_pigs = log(df$pigs2010+addSmall)

#assign nonzero fields
df$pigs_nonzero = 1
df$pigs_nonzero[df$pigs2010==0]=0

df$goats_nonzero = 1
df$goats_nonzero[df$goats2010==0]=0

df$sheep_nonzero = 1
df$sheep_nonzero[df$sheep2010==0]=0

df$cattle_nonzero = 1
df$cattle_nonzero[df$cattle2010==0]=0

df$chickens_nonzero = 1
df$chickens_nonzero[df$chickens2010==0]=0

df$livestock_non_zero_fraction = (df$pigs_nonzero+
                                    df$cattle_nonzero+
                                    df$sheep_nonzero+
                                    df$goats_nonzero+
                                    df$chickens_nonzero)/5

remove_fields = c("chickens2010",
                  "sheep2010",
                  "goats2010",
                  "cattle2010",
                  "pigs2010",
                    "pigs_nonzero",
                    "cattle_nonzero",
                  "sheep_nonzero",
                  "goats_nonzero",
                  "chickens_nonzero")

df$log_highway_density = log(df$highway+addSmall)
df$log_primary_road_density = log(df$primary_roads+addSmall)
df$log_secondary_road_density = log(df$secondary_roads+addSmall)
df$log_tertiary_road_density = log(df$tertiary_roads+addSmall)
df$log_local_road_density = log(df$local_roads+addSmall)
df$logRoadDensity = log(df$RoadDensity+addSmall)

remove_fields = c(remove_fields, "highway",
                  "primary_roads",
                  "secondary_roads",
                  "tertiary_roads",
                  "local_roads",
                  "RoadDensity")

names_df = names(df)
keep_fields = setdiff(names_df, remove_fields)

df = df[,keep_fields]

names(df)
predictorPoints = df
save(predictorPoints, file = "predictorPoints.Rdata")
```

###assign country to predictorPoints
```{r country_predictors1}
load("predictorPoints.Rdata")

r = shapefile("DATA/TM_WORLD_BORDERS-0.3/TM_WORLD_BORDERS-0.3.shp")

out_spdf = predictorPoints
coordinates(out_spdf) <- ~ x + y

#predictorPoints is already a SpatialPointsDataFrame 
projection(r) <-CRS("+proj=longlat +datum=WGS84")
projection(out_spdf) = projection(r)
out_spdf$country <- over(out_spdf, r)$NAME#name of park
#make into dataframe
df= as.data.frame(out_spdf)
keep = names(df)
rm_fields = c("latitude.1", "longitude.1", "ID")
keep = setdiff(keep, rm_fields)
df = df[,keep]

#change to one-hot encoding of country
#Converting every categorical variable to numerical using dummy variables
dmy <- dummyVars(" ~ .", data = df,fullRank = T, sep=".")
df_transformed <- data.frame(predict(dmy, newdata = df))
df = df_transformed

predictorPoints = df
save(predictorPoints, file = "predictorPoints.Rdata")
```

###assign ecoregion to predictorPoints and one-hot encode it
```{r ecoregion_predictors_one_hot}
load("predictorPoints.Rdata")

#source: http://maps.tnc.org/gis_data.html
r = shapefile("ecoregions/tnc_terr_ecoregions.shp")

out_spdf = predictorPoints
coordinates(out_spdf) <- ~ x + y

#predictorPoints is already a SpatialPointsDataFrame 
projection(r) <-CRS("+proj=longlat +datum=WGS84")
projection(out_spdf) = projection(r)
out_spdf$ecoregion <- over(out_spdf, r)$ECO_NAME#name of park
#make into dataframe
df= as.data.frame(out_spdf)
keep = names(df)
rm_fields = c("latitude.1", "longitude.1", "ID")
keep = setdiff(keep, rm_fields)
df = df[,keep]

#change to one-hot encoding of country
#Converting every categorical variable to numerical using dummy variables
dmy <- dummyVars(" ~ .", data = df,fullRank = T, sep=".")
df_transformed <- data.frame(predict(dmy, newdata = df))
df = df_transformed

predictorPoints = df
save(predictorPoints, file = "predictorPoints.Rdata")
```




###construct interaction features: log of the product of roadDensity and livestock density 
```{r interaction_predictors}
load("predictorPoints.Rdata")
df = predictorPoints
df$log_road_chickens = log(exp(df$logRoadDensity)*exp(df$log_chickens))
df$log_road_pigs = log(exp(df$logRoadDensity)*exp(df$log_pigs))
df$log_road_cattle = log(exp(df$logRoadDensity)*exp(df$log_cattle))
df$log_road_sheep = log(exp(df$logRoadDensity)*exp(df$log_sheep))
df$log_road_goats = log(exp(df$logRoadDensity)*exp(df$log_goats))
predictorPoints = df
save(predictorPoints, file = "predictorPoints.Rdata")
```


Add fields to predictorPoints latitude_10 (nearest top 10 degree) and longitude_10 (left 10 degree). Resave predictorPoints.Rdata
```{r deforestation}
source("deforestation_predictors.R")
```

For each 10X10 degree quadrant of outbreak presence / absence data, download relevant tile of deforestation data, and extract deforestation data. Use as input predictorPoints.Rdata. Save out_deforest.  
```{r getdef}
source("getdef_predictors.R")
```

Add field indicating whether location has been deforested or not. Input out_deforest. Make graphs of deforest vs. no for outbreak vs. background locations. Input "predictorPoints"
```{r deforest_binary}
source("deforest_binary_year_predictors.R")
```




###fix column names in predictorPoints to match names in Train
```{r}
load("predictorPoints.Rdata")
group = c(names(predictorPoints))

group = gsub("[.]","", group)
names(predictorPoints)=group
save(predictorPoints, file = "predictorPoints.Rdata")
#check = "country."
# a = 1
# namesOutput = group
# for (a in 1:length(group)){
#   test = c(as.character(names(dfPredictors)[a]))
#   #test = "aa.bb"
#   test_period = strsplit(test, check)
#     if(unique(test_period[[1]])==""){
#  #   if (test_period[[1]][1]==""){
#     namesOutput[a]=test
#     print("empty")
#   } else {
#      test_out = str_replace(test, check, "")
#     namesOutput[a]=test_out   
#   }
  #splitted_count = length(test_period[[1]])
  #test_out= test
  #print(test_out)
  #if (splitted_count>1){
  #  test_out = str_replace_all(test_out, check, "")
    #test_out = gsub(".", "", test_out)
  #}
  #namesOutput[a]=test_out
  
# }
# 
# namesOutput
# 'e' %in% test[1]
# match('_',test)
# 
# is.element('_', test)
# 
# any("livestock_system"=="e")

# test_period = strsplit(test, "_")
# 
# 
# str_replace_all(names(dfPredictors)[1], ".", "")
# group = gsub(".", "", group)
# names(dfPredictors)=group
# names(dfPredictors)

```


###plot predictions
##need to redo this including land cover
```{r predictions_plot}
load("predictorPoints.Rdata")
load("TrainLivestock.Rdata")
names(predictorPoints)[names(predictorPoints)=="label_shortNodataburntareascloudsâ"]="label_shortNodataburntareasclouds"
load("gbmtestLivestock.Rdata")
names_missing = setdiff(names(Train), names(predictorPoints))
names_missing = setdiff(names_missing, "case")
#for each missing name, add a field set to zero for that field
for (a in 1:length(names_missing)){
  predictorPoints[,names_missing[a]]=0
}
#predictorPoints$countryAndorra = 0
best.iter <- gbm.perf(gbmtest,method="cv",plot.it=FALSE) #this gives you the optimal number of trees based on cv 
n.trees = best.iter

sdf = predictorPoints
sdf = subset(sdf, !is.na(log_cattle))
dim(sdf)[1]
sdf = subset(sdf, !is.na(logRoadDensity))
sdf = subset(sdf, livestock_system != 15)
dim(sdf)[1]

sdf_pred = predict.gbm(gbmtest, newdata = sdf, n.trees=n.trees, type = "response")
  sdf$predicted= sdf_pred
  pred_gbm_season = sdf
  pred_gbm_season=data.frame(pred_gbm_season)

#coordinates(pred_gbm_season) <- ~x+y

#world <- map_data("world")
world <- borders("world", ylim = c(-60,80))


mp<- ggplot()+world

    mp <- mp+
   geom_tile(data = pred_gbm_season, aes(x = x, y = y, alpha = predicted))
  ggsave("predictedLivestock.jpg", plot = mp, height = 7, width = 7*2)
  
pred.xy = pred_gbm_season[,c("x", "y", "predicted")]

out_spdf = SpatialPointsDataFrame(coords = pred.xy[,c("x", "y")], 
                               data = pred.xy)

gridded(out_spdf) <- TRUE
out_spdf$z = out_spdf$predicted
dfPredictors_predicted = out_spdf
save(dfPredictors_predicted, file = "dfPredictors_predicted.Rdata")
dfr = raster(out_spdf)

writeRaster(dfr, "prediction_no_naLivestock.tif", format="GTiff", overwrite=TRUE)


```

###load in prediction values and observed locations. Find locations with high predicted probability but no outbreaks  yet observed
```{r get_minimum_distances}
rm(list=ls())
load("dfPredictors_predicted.Rdata")
# load("dfLivestock.Rdata")
load("dfRoad.Rdata")
obs = dfRoad
obs = subset(obs, case == 1)
dfpred = dfPredictors_predicted
dfpred_hi = subset(dfpred, predicted>=0.5)
dim(dfpred_hi)
#https://stackoverflow.com/questions/21977720/r-finding-closest-neighboring-point-and-number-of-neighbors-within-a-given-rad
#coordinates(dfpred_hi) <- ~x+y
coordinates(obs) <- ~longitude+latitude
library(rgeos)
#convert to data frame
dfpred_hi_df = data.frame(dfpred_hi)

dfpred_hi_pts = SpatialPointsDataFrame(dfpred_hi, dfpred_hi_df)

#dfpred_hi_pts = dfpred_hi_pts[1,]

#Start the clock

nrows = dim(dfpred_hi_df)[1]
distances = rep(NA, nrows)
a = 1
a = 2
ptm<-proc.time()

#for (a in 1){
for (a in 1:nrows){
  print(a)
  tmp = dfpred_hi_pts[a,]
  distances[a] = gDistance(tmp, obs, byid=FALSE)#specify byid=FALSE in order
  # to get the minimum distance
}
 
#Stop the clock
(proc.time()-ptm)/60

#save(distances, file = "distances.Rdata")
dfpred_hi_pts$minimum_distance = distances
save(dfpred_hi_pts, file = "dfpred_hi_pts.Rdata")


```



###SCRATCH below here.

###format date for python
```{r}

# load("dfRoads.Rdata")
# df = dfRoads
# is.na.date = which(is.na(df$date))
# #fix date that has weird format
# df$date[is.na.date]= as.Date(df$reportingDate[is.na.date], format = "%d/%m/%Y")
# df$month = month(df$date)
# is.na.month = which(is.na(df$month))#should be empty
# 
# #beginning of the year
# df$year_begin = paste0(df$year,"0101")
# #beginning of the previous year
# df$previous_year_begin = paste0(df$year-1,"0101")
# #next year beginning
# df$next_year_begin = paste0(df$year+1,"0101")
# dfDate = df
# save(dfDate, file = "dfDate.Rdata")

```

```{r}
#https://trucvietle.me/r/tutorial/2017/01/18/spatial-heat-map-plotting-using-r.html
# m <- ggplot(faithful, aes(x = eruptions, y = waiting)) +
#  geom_point() +
#  xlim(0.5, 6) +
#  ylim(40, 110)
# m + geom_density_2d()
# m+ stat_density_2d(aes(fill = stat(level)), geom = "polygon")

```

###plot density 
```{r}

# load("dfRoads.Rdata")
# df = dfRoads
# df = df[,c("latitude", "longitude")]
# m <- ggplot(df, aes(x = longitude, y = latitude)) +
#  geom_point() 
#  # xlim(0.5, 6) +
#  # ylim(40, 110)
# m + geom_density_2d()
# m+ stat_density_2d(aes(fill = stat(level)), geom = "polygon")

```

####interpolation with kriging
###https://rpubs.com/nabilabd/118172
```{r}

# load("dfRoads.Rdata")
# df = dfRoads
# library(sp)
# library(gstat)
# suppressPackageStartupMessages({
#   library(dplyr) # for "glimpse"
#   library(ggplot2)
#   library(scales) # for "comma"
#   library(magrittr)
# })
# 
# #1) Converting to an SPDF
# coordinates(df) <- ~ longitude + latitude
# #2) Fitting a variogram
# lzn.vgm <- variogram(case~1, df) # calculates sample variogram values 
# lzn.fit <- fit.variogram(lzn.vgm, model=vgm(1, "Sph", 900, 1)) # fit model

```

